---
title: "project"
author: "Dehshini"
date: "2023-12-28"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Regression Model Course Project

Executive summary:

This analysis was done to answer the following:

“Is an automatic or manual transmission better for MPG”?\
"Quantify the MPG difference between automatic and manual transmissions"

There was a total of 19 automatic transmission cars and 13 manual transmission cars. The average mpg for automatic transmission cars was 17.15mpg while that for manual cars was higher by 7.25 at 24.39mpg. After controlling for factors that may increase fuel consumption such as car weight, number of cylinders and gross horsepower of engines, there was no significant difference in mpg between automatic and manual transmission cars. The difference in mpg between manual and automatic was 1.8 (95% CI -1.1, 4.7)

Conclusion: there is a 1.8mpg difference between manual and automatic transmission cars after controlling for car weight, gross horsepower and number of cylinders of engine. Both automatic and manual transmission have similar mpg.

```{r look, echo=FALSE}
mt = mtcars
head(mt)
tail(mt)
summary(mt)

mt %>% 
  group_by(am) %>% 
  summarise(
    n = n(),
    mean_mpg = mean(mpg)
  )
```

```{r view, include=FALSE}
boxplot(mpg~factor(am), data = mt)
```

mpg seems to be associated with transmission with the automatic transmission cars having less mpg compared to manual transmission cars.

```{r regression, echo=FALSE}
reg_s = lm(mpg~factor(am), data=mt)
summary(reg_s)
confint(reg_s)
```

the average mpg of automatic transmission cars is 17.147 (95% confidence interval 14.8 - 19.4). the mean mpg of manual transmission vehicles is 7.245 more than automatic transmission vehicles(95% CI 3.6 - 10.8).

this relationship may be confounded by other variables such as weight, horsepower, number of cylinders. Do a matrix graph to explore the associations.

control for other variables

```{r echo=FALSE}
reg_m1 = lm(mpg~factor(am)+factor(cyl)+wt+hp, data=mt)
summary(reg_m1)
confint(reg_m1)
```

```{r include=FALSE}
#control for all variables
reg_m2 = lm(mpg~., data=mt)
summary(reg_m2)
anova(reg_m2)
```

after controlling for weight of car, number of cylinders, and gross horsepower, the mpg does not significantly differ by transmission type. the average mpg for automatic transmission cars is 33.7 (95% CI 28.4 - 39.1). the average for manual transmission is 1.8 more than that of automatic (95% CI -1.1 - 4.7)

compare the two models

```{r anova, include=FALSE}
anova(reg_s, reg_m1)
anova(reg_m1, reg_m2)
```

the multivariate model seems better as the residual sum of squares is significantly lower.

check model assumptions.

```{r resid_v_fitted, echo=FALSE}
plot(x=reg_m1$fitted.values, y=reg_m1$residuals)
abline(h=0, col="red")

```

residuals look random with no clear pattern. Linear, somewhat equal variance except for the higher fitted values where variance is higher.

```{r include=FALSE}
plot(reg_m1)
```

```{r qq, include=FALSE}
qqnorm(reg_m1$residuals)
qqline(reg_m1$residuals)
```

normal distribution mostly, except for the tail ends.

perform diagnostic, dfits

```{r include=FALSE}
#dfbetas(reg_m1)
dffits(reg_m1)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
pairs(~ wt + hp + factor(cyl) + factor(am) + mpg, data=mt, upper.panel=NULL, pch=".")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
